---
title: "Descriptives"
author: "Caroline Ledbetter"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3.5, fig.height=3, 
                      echo=FALSE, warning=FALSE, message=FALSE)
```
```{r setup, include=FALSE}
load(file = "~/Repositories/bios6623-ledbettc/Project1/Descriptives.RData")
library(pander)
```

There were `r length(camp_teach$id) ` observations for `r length(ids)` unique subjects. The number of observations for each subject ranged from `r min(ids_obs$obs)` to `r max(ids_obs$obs)` with an average of `r sprintf('%.2f',mean(ids_obs$obs))` observations (Fig 1). Charachteristics for study participants are given in Table 1. To confirm random assignment, categorical variables were tested using $\chi ^2$ and continuous variables were tested using ANOVA. p-values are shown in Table 1.

```{r table one}
panderOptions('table.split.table', Inf)
set.alignment('center', row.names = 'left')
pander(Table_1, caption = 'Characteristics of Study Participants by Treamtment Group')
```

```{r hist of obs}
hist(ids_obs$obs, freq = TRUE, plot = TRUE, xlab= "Number of Observations per Subject", col = "blue",
     main='Frequency of Observations
     per Subject', breaks = c((min(ids_obs$obs - 0.5)):(max(ids_obs$obs + 0.5))), xlim = c(0,20), cex.main = 1, cex.lab = 0.75)
```

```{r hist of times}
hist(camp_teach$visitc, freq = TRUE, plot = TRUE, xlab= 
       "Months from Initial Measurement", col = "blue",
     main='Number of Ids with observations
at each time period', cex.main = 1, cex.lab = 0.75)
```