---
title: "Project0_Descriptives"
author: "Caroline Ledbetter"
date: "August 30, 2016"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
load(file = '/Users/CarolineL/Repositories/Data/Project0Data/Project0cleandata.rda')
library(pander)
library(doBy)
```
A preliminary analysis of the dental data set. 
There were `r nrow(Project0_dental_data)` observations in the data set. There were `r nrow(Project0_dental_data[is.na(Project0_dental_data$attach1year),])` observations missing final measurements. `r nrow(Project0_dental_data[is.na(Project0_dental_data$attach1year) & Project0_dental_data$trtgroup == 5,])` of these missing observations were in the high treatment group. (Table 1.)
```{r table 1}
pander(summaryBy(attach1year + pd1year ~ trtgroup, data = Project0_dental_data,
          FUN=function(x)sum(is.na(x)),keep.names = TRUE), caption = "Number of Missing Values by Treatment Group")
```


```{r tables}
pander(table(Project0_dental_data$sex), caption = "Frequency of Gender")
pander(table(Project0_dental_data$race), caption = "Frequency of Race")
pander(table(Project0_dental_data$smoker), caption = "Frequency of Smoking Status")
```

```{r figures}
barplot(table( Project0_dental_data$race,Project0_dental_data$trtgroup),
        col = c("blue", "green","red", "light blue"), 
        main = "Race by Treatment Group", legend.text = T,
        args.legend = list(bty = 'n', cex = 2.5))

barplot(table( Project0_dental_data$sex,Project0_dental_data$trtgroup),
        col = c("blue", "red"), 
        main = "Sex by Treatment Group", legend.text = T,
        args.legend = list(bty = 'n', cex = 2.5))

boxplot(Project0_dental_data$age~Project0_dental_data$trtgroup,
        main = "Age by Treatment Group",
        xlab = 'Treatment Group', ylab = 'Age')
```

__A preliminary look at results.__

```{r box plot results}
boxplot(Project0_dental_data$attachdiff ~ Project0_dental_data$trtgroup,
        main = 'Difference in attachment measurements
        at 1 year by Treatment Group',
        ylab = 'Attachment Difference',
        xlab = 'Treatment Group')
boxplot(Project0_dental_data$attachdiff[Project0_dental_data$smoker == 'No']
        ~ Project0_dental_data$trtgroup[Project0_dental_data$smoker == 'No'],
        main = 'Difference in attachment measurements
        at 1 year by Treatment Group
        for Non Smokers',
        ylab = 'Attachment Difference',
        xlab = 'Treatment Group')
boxplot(Project0_dental_data$attachdiff[Project0_dental_data$smoker == 'Yes']
        ~ Project0_dental_data$trtgroup[Project0_dental_data$smoker == 'Yes'],
        main = 'Difference in attachment measurements
        at 1 year by Treatment Group
        for Smokers',
        ylab = 'Attachment Difference',
        xlab = 'Treatment Group')
```
