---
title: "Comprehensive Analysis"
author: "Caroline Ledbetter"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3.5, fig.height=3, 
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
load(file = '/Users/Caroline/Repositories/Data/Project0Data/Project0cleandatanomissing.rda')
attach(Project0_no_missing)
library(pander)
library(doBy)
```
This is the comprehensive analysis report for the Project 0 Dental Data. This study is interested in the effect of a gel treatment on gum health as measured by attachment loss and pocket depth. Measurements for both outcomes were taken at a wide variety of sites and averaged to determine a whole mouth value. Measurements were collected at baseline before treatment began and after one-year of treatment. The study participants were randomly assigned to one of five-treatment groups -  no-treatment (no-gel), gel only (no-active ingredient/placebo), low dose, medium dose, high dose. The characteristics of each group are shown in Table 1. To confirm random assignment, categorical variables were tested using $\chi ^2$ and Fisher's Exact and continuous variables were tested using ANOVA. p-values are shown in Table 1. 
```{r table1}
source(file = '/Users/Caroline/Repositories/Table1/Table1.R')
rowvars <- c(3,5:8,10,15)
library(Hmisc)
names(Project0_no_missing)[c(3,5,6)] <- lapply(c(3,5,6), function(i)capitalize(names(Project0_no_missing)[i]))
names(Project0_no_missing)[c(7,8,10,15)] <- c("No of Sites", "Baseline Attachment", "Baseline Pocket Depth", "Race")

Table_1 <- Table1(rowvars,12,Project0_no_missing)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', 20)
set.alignment('center', row.names = 'left')
pander(Table_1, caption = 'Characteristics of Study Participants by Treamtment Group')
```

The only potential covariate with a significant relationship to the exposure was attachment loss at baseline. This would also be expected to effect outcome and was therefore adjusted for in the final model.

```{r function for p-values}
p_value <- function(model){
  model_summary <- summary(model)
  p_value <- pf(model_summary$fstatistic[1],model_summary$fstatistic[2],model_summary$fstatistic[3])
  return(p_value)
}
```

```{r models}
#####attachment#####
#crude
logattach <- log(attach1year)
lmlogattach <- lm(logattach~trtgroupFactored)
p_crude_attach <- p_value(lmlogattach)
#adjusted 
reduced_attach_final <- lm(attach1year ~ attachbase + sex +  smoker + sites)
full_attach_final <- lm(attach1year ~ attachbase + sex +  smoker + sites + trtgroupFactored)
adjusted_attach <- anova(reduced_attach_final, full_attach_final)

####pocket depth#####
#crude
logpd <- log(pd1year)
lmlogpd <- lm(logpd~trtgroupFactored)
p_crude_pd <- p_value(lmlogpd)
#adjusted
reduced_pd2 <- lm(pd1year ~ pdbase + sites + smoker)
full_pd2 <- lm(pd1year ~ pdbase + sites + smoker + trtgroupFactored) 
adjusted_pd <- anova(reduced_pd2, full_pd2)

```
__Attachment Loss__ \newline
Due to problems with normality and homoscedasticity, the outcome was log transformed for the crude analysis. There was a significant difference in treatment groups when looking at attachment loss (p=`r round(p_crude_attach, digits = 4)`), but further analysis showed the only significant difference from no-treatment was with the placebo group (Table 2). There was no significant difference between treatment groups after adjusting for baseline, number of sites used, smoking status and gender (p=`r round(adjusted_attach[2,"Pr(>F)"],digits = 4)`).
\newpage
```{r Table2 crude univariate analysis}
  rn <- c("Intercept", levels(trtgroupFactored)[-1])
  cn <- c("Estimate", "t value", "Pr(>|t|")
  table_2 <- summary(lmlogattach)$coefficients[,c(1,3,4)]
  dimnames(table_2) <- list(rn,cn)
  pander(table_2, caption = "Crude Model of Attachment Loss")
```

__Pocket Depth__\newline
Due to problems with normality and homoscedasticity, the outcome was log transformed for the crude analysis. There was no significant difference in treatment groups when looking at attachment loss (p=`r round(p_crude_pd, digits = 4)`). There was no significant difference between treatment groups after adjusting for baseline, number of sites used, and smoking status (p=`r round(adjusted_pd[2,"Pr(>F)"],digits = 4)`).

__Appendix: Diagnostic Plots__\newline
```{r plot setup}
library(car)
#fxn for histogram of studentized residuals
hist_resid <- function(lm){
  require(MASS)
  sresid <- studres(lm) 
  hist(sresid, freq=FALSE, main = "Histogram of Residuals", xlab = 'Residuals')
  xfit<-seq(min(sresid),max(sresid),length=40) 
  yfit<-dnorm(xfit) 
  lines(xfit, yfit)
}
```

$\textit{Diagnostic Plots of Attachment Loss vs Treatment Group}$\newline
\newline
```{r linear attach plots}
lm1yrattach <-lm(attach1year~trtgroupFactored)
qqPlot(lm1yrattach, main = "QQ Plot", ylab = 'Studentized Residuals')
hist_resid(lm1yrattach)
```

$\textit{Diagnostic Plots of Log Transformed Attachment Loss vs Treatment Group}$\newline
\newline
```{r log attach plots}
qqPlot(lmlogattach, main = "QQ Plot", ylab = 'Studentized Residuals') 
hist_resid(lmlogattach)
```

$\textit{Diagnostic Plots of Pocket Depth vs Treatment Group}$\newline
\newline
```{r linear pocket depth plots}
lm1yrpd <- lm(pd1year~trtgroupFactored)
qqPlot(lm1yrpd, main = "QQ Plot", ylab = 'Studentized Residuals')
hist_resid(lm1yrpd)
```

$\textit{Diagnostic Plots of Log Transformed Pocket Depth vs Treatment Group}$\newline
\newline
```{r log pocket depth plots}
qqPlot(lmlogpd, main = "QQ Plot", ylab = 'Studentized Residuals') 
hist_resid(lmlogpd)
```