---
title: "Preliminary Analysis"
author: "Caroline Ledbetter"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
load(file = '/Users/Caroline/Repositories/Data/Project0Data/Project0cleandatanomissing.rda')
library(pander)
library(doBy)
```
This is the prelimianry analysis report for the Project 0 Dental Data. This study is interested in the effect of a gel treatment on gum health as measured by attachment loss and pocket depth. Measurements for both outcomes were taken at a wide variety of sites and averaged to determine a whole mouth value. Measurements were collected a baseline before treatment began and after one-year of treatment. The study participants were randomly assigned to one of five-treatment groups -  no-treatment (no-gel), gel only (no-active ingredient/placebo), low dose, medium dose, high dose. The charactersitics of each group are shown in Table 1. To confirm random assignment, categorical variables were tested using $\chi ^2$ and Fisher's Exact and continuous variables were tested using ANOVA. p-values are shown in Table 1. 
```{r table1}
source(file = '/Users/Caroline/Repositories/Table1/Table1.R')
rowvars <- c(3,5:8,10,15)
library(Hmisc)
names(Project0_no_missing)[c(3,5,6)] <- lapply(c(3,5,6), function(i)capitalize(names(Project0_no_missing)[i]))
names(Project0_no_missing)[c(7,8,10,15)] <- c("No of Sites", "Baseline Attachment", "Baseline Pocket Depth", "Race")

Table_1 <- Table1(rowvars,12,Project0_no_missing)

panderOptions('table.split.table', 300)
set.alignment('center', row.names = 'left')
pander(Table_1, caption = 'Characteristics of Study Participants by Treamtment Group')
```
```{r Table2 univariate analysis}
attach(Project0_no_missing)
lm1yrattach_summary <-summary(lm(attach1year~trtgroupFactored))
lm1yrpd_summary <-summary(lm(pd1year~trtgroupFactored))
pander(lm1yrattach_summary, caption = "Results of linear regression model of difference in mean attachment loss at 1-year by treatment group")
```
The only potential covariate with a signifiant relationship to the exposure was attachment loss at baseline. This would also be expected to effect outcome and was therefore adjusted for in the final model.
The overall p-value for the univariate model comparing mean attachment loss at 1-year was `r round(pf(lm1yrattach_summary$fstatistic[1],lm1yrattach_summary$fstatistic[2],lm1yrattach_summary$fstatistic[3], lower.tail = F),digits = 4)`. For the univariate model comparing mean pocket depth at 1 -year was `r round(pf(lm1yrpd_summary$fstatistic[1],lm1yrpd_summary$fstatistic[2],lm1yrpd_summary$fstatistic[3], lower.tail = F), digits = 4)`. As you can see from table 2, pairwise t-test show the significant difference in attachment loss at 1-year from the no treatment group is the placebo group. 

Additionaly, because attachment loss at baseline was significantly different across treatment groups, a linear regression adjusted for baseline measurements was also preformed. 

``` {r table 3djusted attachment model}
lm1yrattach_base_summary <- summary(lm(attach1year ~ trtgroupFactored + Project0_no_missing[,8]))
pander(lm1yrattach_base_summary, caption = "Results of linear regression model of difference in mean attachment loss at 1-year by treatment group adjusted for baseline measurement")
```
When adjusting for baseline attachment, there was a significant difference between the medium treatment and no treatment (Table 3).
